---
title: "Multi-Hospital Branding"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multi-Hospital Branding}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

BFHtheme supports branding for both Bispebjerg og Frederiksberg Hospital and Region Hovedstaden (koncern level). This vignette demonstrates how to work with different organizational branding requirements.

```{r load, message=FALSE}
library(BFHtheme)
library(ggplot2)
```

## Organizational Color Palettes

### Bispebjerg og Frederiksberg Hospital

The hospital-level branding uses blue as the primary color:

```{r hospital-colors}
# Access hospital colors
hospital_primary <- bfh_cols("hospital_primary")
hospital_blue <- bfh_cols("hospital_blue")

print(hospital_primary)
print(hospital_blue)

# View all hospital colors
hospital_palette <- bfh_cols(
  "hospital_primary",
  "hospital_blue",
  "hospital_light_blue1",
  "hospital_light_blue2",
  "hospital_grey"
)
print(hospital_palette)
```

Use hospital palettes in plots:

```{r hospital-plot}
ggplot(mtcars, aes(x = factor(cyl), fill = factor(cyl))) +
  geom_bar() +
  labs(
    title = "Hospital Branding Example",
    x = "Cylinders",
    fill = "Cylinders"
  ) +
  scale_fill_bfh(palette = "hospital") +
  theme_bfh()
```

### Region Hovedstaden

The regional branding uses navy blue as the primary color:

```{r regionh-colors}
# Access Region H colors
regionh_primary <- bfh_cols("regionh_primary")
regionh_blue <- bfh_cols("regionh_blue")

print(regionh_primary)
print(regionh_blue)

# View all Region H colors
regionh_palette <- bfh_cols(
  "regionh_primary",
  "regionh_blue",
  "regionh_light_grey1",
  "regionh_light_grey2",
  "regionh_grey"
)
print(regionh_palette)
```

Use Region H palettes in plots:

```{r regionh-plot}
ggplot(mtcars, aes(x = factor(cyl), fill = factor(cyl))) +
  geom_bar() +
  labs(
    title = "Region Hovedstaden Branding",
    x = "Cylinders",
    fill = "Cylinders"
  ) +
  scale_fill_bfh(palette = "regionh") +
  theme_bfh()
```

## Sequential Palettes

### Hospital Sequential Palettes

For data visualizations requiring sequential color scales:

```{r hospital-sequential}
# Create sample data
set.seed(123)
data <- expand.grid(x = 1:10, y = 1:10)
data$z <- with(data, x + y + rnorm(100, 0, 2))

ggplot(data, aes(x = x, y = y, fill = z)) +
  geom_tile() +
  labs(title = "Hospital Blues Sequential Palette") +
  scale_fill_bfh_continuous(palette = "hospital_blues_seq") +
  theme_bfh()
```

### Region H Sequential Palettes

```{r regionh-sequential}
ggplot(data, aes(x = x, y = y, fill = z)) +
  geom_tile() +
  labs(title = "Region H Blues Sequential Palette") +
  scale_fill_bfh_continuous(palette = "regionh_blues_seq") +
  theme_bfh()
```

## Infographic Palettes

For presentations and infographics with multiple distinct categories:

### Hospital Infographic Palette

```{r hospital-infographic}
# Sample categorical data
category_data <- data.frame(
  category = LETTERS[1:5],
  value = c(23, 45, 12, 34, 28)
)

ggplot(category_data, aes(x = reorder(category, value), y = value, fill = category)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Hospital Infographic Colors",
    x = "Category",
    y = "Value"
  ) +
  scale_fill_bfh(palette = "hospital_infographic") +
  theme_bfh() +
  theme(legend.position = "none")
```

### Region H Infographic Palette

```{r regionh-infographic}
ggplot(category_data, aes(x = reorder(category, value), y = value, fill = category)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Region H Infographic Colors",
    x = "Category",
    y = "Value"
  ) +
  scale_fill_bfh(palette = "regionh_infographic") +
  theme_bfh() +
  theme(legend.position = "none")
```

## Combining Organizational Branding

### Consistent Theme with Different Palettes

Use the same theme structure across organizations with different color schemes:

```{r combined-hospital}
# Hospital version
p_hospital <- ggplot(mtcars, aes(x = wt, y = mpg, color = hp)) +
  geom_point(size = 3) +
  labs(
    title = "Hospital Analysis",
    subtitle = "Vehicle Performance Data"
  ) +
  scale_color_bfh_continuous(palette = "hospital_blues") +
  theme_bfh()

print(p_hospital)
```

```{r combined-regionh}
# Region H version
p_regionh <- ggplot(mtcars, aes(x = wt, y = mpg, color = hp)) +
  geom_point(size = 3) +
  labs(
    title = "Region Hovedstaden Analysis",
    subtitle = "Vehicle Performance Data"
  ) +
  scale_color_bfh_continuous(palette = "regionh_blues") +
  theme_bfh()

print(p_regionh)
```

### Setting Default Palette

Set organizational defaults for a session:

```{r defaults-hospital, eval=FALSE}
# Set hospital as default
set_bfh_defaults(palette = "hospital")

# All subsequent plots use hospital colors
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar()
```

```{r defaults-regionh, eval=FALSE}
# Switch to Region H defaults
set_bfh_defaults(palette = "regionh")

# All subsequent plots use Region H colors
ggplot(mtcars, aes(x = factor(cyl))) +
  geom_bar()
```

## Color Accessibility

### Contrast and Readability

Both hospital and Region H palettes are designed for adequate contrast:

```{r contrast}
# Hospital primary on white background
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = bfh_cols("hospital_primary"), size = 3) +
  labs(title = "Hospital Primary Color") +
  theme_bfh()

# Region H primary on white background
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_point(color = bfh_cols("regionh_primary"), size = 3) +
  labs(title = "Region H Primary Color") +
  theme_bfh()
```

### Colorblind Considerations

Check palette accessibility:

```{r colorblind, eval=FALSE}
# Basic check
check_colorblind_safe(bfh_cols("hospital_primary", "hospital_blue"))

# For comprehensive testing, use specialized packages
# install.packages("colorblindcheck")
library(colorblindcheck)

palette_check(
  bfh_palettes$hospital,
  plot = TRUE
)
```

## Customizing for Specific Departments

### Department-Specific Colors

Extend the color system for department use:

```{r department-colors}
# Define department colors using base palette
dept_emergency <- bfh_cols("hospital_primary")
dept_surgery <- bfh_cols("hospital_blue")
dept_pediatrics <- bfh_cols("hospital_light_blue1")

dept_colors <- c(
  "Emergency" = dept_emergency,
  "Surgery" = dept_surgery,
  "Pediatrics" = dept_pediatrics
)

# Use in plots
dept_data <- data.frame(
  department = names(dept_colors),
  patients = c(120, 85, 95)
)

ggplot(dept_data, aes(x = department, y = patients, fill = department)) +
  geom_col() +
  scale_fill_manual(values = dept_colors) +
  labs(
    title = "Patient Count by Department",
    x = "Department",
    y = "Number of Patients"
  ) +
  theme_bfh() +
  theme(legend.position = "none")
```

## Multi-Site Comparisons

### Comparing Data Across Sites

Maintain consistent styling when comparing multiple sites:

```{r multi-site}
# Sample multi-site data
site_data <- data.frame(
  site = rep(c("BFH", "Rigshospitalet", "Hvidovre"), each = 4),
  quarter = rep(paste0("Q", 1:4), 3),
  value = c(
    85, 88, 92, 90,  # BFH
    78, 82, 85, 83,  # Rigshospitalet
    82, 85, 88, 86   # Hvidovre
  )
)

ggplot(site_data, aes(x = quarter, y = value, group = site, color = site)) +
  geom_line(linewidth = 1) +
  geom_point(size = 3) +
  labs(
    title = "Quarterly Performance Across Sites",
    x = "Quarter",
    y = "Performance Score",
    color = "Site"
  ) +
  scale_color_manual(
    values = c(
      "BFH" = bfh_cols("hospital_primary"),
      "Rigshospitalet" = bfh_cols("regionh_primary"),
      "Hvidovre" = bfh_cols("hospital_grey")
    )
  ) +
  theme_bfh()
```

## Brand Guidelines Reference

### Hospital Colors

**Primary**: `hospital_primary` (#007dbb)
- Use: Main brand color, primary data series
- Usage: Headers, primary chart elements

**Secondary**: `hospital_blue` (#009ce8)
- Use: Secondary data series, accents
- Usage: Supporting chart elements

**Light Blues**: `hospital_light_blue1`, `hospital_light_blue2`
- Use: Backgrounds, fills, lighter emphasis
- Usage: Sequential scales, area fills

**Neutrals**: `hospital_grey`, `hospital_dark_grey`
- Use: Text, borders, secondary information
- Usage: Axis labels, grid lines

### Region Hovedstaden Colors

**Primary**: `regionh_primary` (#002555)
- Use: Main brand color for regional reporting
- Usage: Headers, primary emphasis

**Secondary**: `regionh_blue` (#007dbb)
- Use: Secondary series, complements primary
- Usage: Supporting elements

**Light Greys**: `regionh_light_grey1`, `regionh_light_grey2`
- Use: Backgrounds, subtle fills
- Usage: Sequential scales, backgrounds

**Neutrals**: `regionh_grey`, `regionh_dark_grey`
- Use: Text, neutral elements
- Usage: Labels, annotations

## Usage Guidelines

### When to Use Hospital vs Region H Branding

**Hospital Branding**:
- Internal BFH reports and presentations
- Department-level analyses
- Hospital-specific communications
- Patient-facing materials for BFH

**Region H Branding**:
- Regional reports and comparisons
- Cross-hospital analyses
- Regional management presentations
- Koncern-level communications

### Consistency Best Practices

1. Use one primary palette per document
2. Maintain consistent palette choice across related documents
3. Use infographic palettes for categorical data with many groups
4. Use sequential palettes for continuous data
5. Test color combinations for accessibility
6. Document palette choice in analysis scripts

## Next Steps

- See `vignette("getting-started")` for basic usage patterns
- See `vignette("customization")` for advanced options
- Review organizational brand guidelines for detailed specifications
- Check `?bfh_cols` and `?bfh_palettes` for complete color reference
